Paper Analysis

### 2025-04-30 Paper_WriteUp/Paper_Plots.qmd
implementing RDs in parallel

#plotting RDs and volatilities for reviewer response
```{r}
#import data
data_path = paste0(all_data_path[1], "/")

@@

start = Sys.time()
init_data <- tidy_games |>
  mutate(WhiteElo = as.numeric(WhiteElo),
         BlackElo = as.numeric(BlackElo),
         focal_user = ifelse(focal_white == 1, White, Black),
         elo_diff = ifelse(focal_white == 1,
                           WhiteElo - BlackElo, BlackElo - WhiteElo),
         focal_id = match(focal_user, users),
         focal_rating = ifelse(focal_white == 1, WhiteElo, BlackElo),
         opp = ifelse(focal_white == 1, Black, White),
         opp_rating = ifelse(focal_white == 1, BlackElo, WhiteElo),
         UTCDateTime = ymd_hms(paste0(UTCDate, "_", UTCTime))) |>
  dplyr::select(focal_user, focal_id, focal_white, WhiteElo, BlackElo, White, Black,
                focal_win_prop, elo_diff, focal_result, opp, opp_rating,
                focal_rating, UTCDateTime) |>
  group_by(focal_id) |>
  mutate(time_diff = UTCDateTime - lag(UTCDateTime, default = NA), #default is to ensure first game is always start of a new session
         cum_win_prob = cummean(focal_result), #the mean win probability for the focal player up to the ith (current) game 
         ave_prop = ifelse(time_diff > 300 | is.na(time_diff),  
                           cum_win_prob, #if games played in different session, history is their mean win prob up to the current game
                           lag(focal_win_prop))) |> #if game played in same session, rolling mean over the past n games, removing standardization 
  filter(focal_result != 0.5) %>%
  ungroup()

#get RDs
init_RDs = future_map_dfr(users, get_RDs, init_data) |>
  as_tibble()

end = Sys.time()
end - start #takes like 20 minutes

#RDs
RD_plot = ggplot(data = init_RDs %>% group_by(focal_user) %>% mutate(index = row_number()), mapping = aes(x = index, y = RD)) +
  geom_line() +
  scale_x_continuous(n.breaks = 3) +
  facet_wrap(~focal_user)

@@

#volatilities
vol_plot = ggplot(data = init_RDs %>% group_by(focal_user) %>% mutate(index = row_number()), mapping = aes(x = index, y = volatility)) +
  geom_line() +
  scale_x_continuous(n.breaks = 3) +
  facet_wrap(~focal_user)

>>>>

#plotting RDs and volatilities for reviewer response
```{r}
#import data
data_path = paste0(all_data_path[4])

files <- list.files(data_path)

lichess_data <- files |> 
  map_dfr(~read_player(paste0(data_path, "/"), .x))
users <- readRDS(file = here("results_revision", "lichessGrandmasters", "users_bullet.RDS"))

small_data <- lichess_data |>
  mutate(Event = tolower(Event)) |>
  filter(Event == paste0("rated bullet game"),
         Variant == "Standard") |>
  filter(TimeControl %in% c("60+0", "180+0")) |> 
  distinct() #remove the duplicate rows if they exist

@@

tidy_games <- map_dfr(users, get_hist, small_data, prev_n = 1) |> 
  as_tibble()

init_data <- tidy_games |>
  mutate(WhiteElo = as.numeric(WhiteElo),
         BlackElo = as.numeric(BlackElo),
         focal_user = ifelse(focal_white == 1, White, Black),
         elo_diff = ifelse(focal_white == 1,
                           WhiteElo - BlackElo, BlackElo - WhiteElo),
         focal_id = match(focal_user, users),
         focal_rating = ifelse(focal_white == 1, WhiteElo, BlackElo),
         opp = ifelse(focal_white == 1, Black, White),
         opp_rating = ifelse(focal_white == 1, BlackElo, WhiteElo),
         UTCDateTime = ymd_hms(paste0(UTCDate, "_", UTCTime))) |>
  dplyr::select(focal_user, focal_id, focal_white, WhiteElo, BlackElo, White, Black,
                focal_win_prop, elo_diff, focal_result, opp, opp_rating,
                focal_rating, UTCDateTime) |>
  group_by(focal_id) |>
  mutate(time_diff = UTCDateTime - lag(UTCDateTime, default = NA), #default is to ensure first game is always start of a new session
         cum_win_prob = cummean(focal_result), #the mean win probability for the focal player up to the ith (current) game 
         ave_prop = ifelse(time_diff > 300 | is.na(time_diff),  
                           0, #if games played in different session, history is their mean win prob up to the current game
                           lag(focal_win_prop) - cum_win_prob)) |> #if game played in same session, rolling mean over the past n games, removing standardization 
  filter(focal_result != 0.5) %>%
  ungroup()

#get RDs
plan(multisession, workers = 12) #in parallel - dramatically speeds it up
start = Sys.time()
init_RDs = future_map_dfr(users, get_RDs, init_data) |>
  as_tibble()
end = Sys.time()
end - start #takes 1.3 min

#RDs
RD_plot = ggplot(data = init_RDs %>% group_by(focal_user) %>% mutate(index = row_number()), mapping = aes(x = index, y = RD)) +
  geom_line() +
  scale_x_continuous(n.breaks = 3, limits = c(0, 10000)) +
  labs(x = "Game Number", y = "Rating Deviation") +
  facet_wrap(~focal_user)

@@

#volatilities
vol_plot = ggplot(data = init_RDs %>% group_by(focal_user) %>% mutate(index = row_number()), mapping = aes(x = index, y = volatility)) +
  geom_line() +
  scale_x_continuous(n.breaks = 3, limits = c(0, 10000)) +
  labs(x = "Game Number", y = "Rating Deviation") +
  facet_wrap(~focal_user)
